---
title: "titanic decision tree"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tree)
library(tidyverse)
```

1. Loading dataset
```{r}
data <- read.csv("titanic_train.csv")
data$Survived <- factor(data$Survived)
```
2. Calculating percentage of passengers survived
```{r}
as.numeric(levels(data$Survived)[data$Survived]) %>% mean()
```
3. choosing variables

I think sex, age, sibsp,parch, pclass and ticket are all good predictors.
sex and age are good predictors since they often let children and women leave first.
sibsp,parch are important since family often will try to save each other.(e.g. the parents will try hard to save their children first and etc)
pclass and ticket are also good predictors since I think the upperclass will be allowed to leave first as well

4.descision tree with age and sex

```{r}
tree.sex_age <- tree(Survived~Age + Sex, data)
summary(tree.sex_age)
```
5. descision tree with age and sex and pclass
```{r}
tree.sex_age_pclass <- tree(Survived~Age + Sex + as.factor(Pclass), data)
summary(tree.sex_age_pclass)
#it has a better error rate than the tree with just age and sex
```

6. descision tree with age and sex and pclass and parch sibsp
```{R}
data$Pclass <- factor(data$Pclass)
tree.sex_age_pclass_parch_sibsp <- tree(Survived ~ Age + Sex + Pclass + SibSp, data)
summary(tree.sex_age_pclass_parch_sibsp)

plot(tree.sex_age_pclass_parch_sibsp)
text(tree.sex_age_pclass_parch_sibsp,pretty=0)
#slight decrease in error rate
```
7. loading test set and predicting with it
```{r}
test <- read.csv("titanic_test.csv")
test$Pclass <- factor(test$Pclass)
pred <- predict(tree.sex_age_pclass_parch_sibsp, newdata = test, type = "class")
submit <- data.frame(PassengerId=test$PassengerId,Survived =pred)
write.csv(submit,"submission.csv", row.names=F, quote = F)
```

```{r}
data$Survived_Raw <- as.numeric(levels(data$Survived)[data$Survived])
data$age_bin <- cut(data$Age, breaks = seq(0,80,by = 5))

data$married <- grepl("Mrs",data$Name)
data[grep("Miss",data$Name),]$Survived_Raw %>%  mean()
data %>% filter(!married)%>% .$Survived_Raw %>% mean()
data[grep("Mrs",data$Name),]$Survived_Raw %>% mean()
data%>% filter(SibSp == 1) %>% .$Survived_Raw %>% mean()

data %>% filter(grep("Miss", Name))
data %>% group_by(Embarked)%>% summarize(rate = mean(Survived_Raw), count = n()) %>% 
ggplot(aes(x=Embarked, y=rate)) + geom_point()
```