---
title: "hw3"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse);
oj <- read.csv("oj.csv")
```

1. Let's return to the orange juice dataset and investigate how store demographics are related to demand.
    a. **Take the "fully interacted" model from HW2 `logmove ~ log(price)\*brand\*feat` and add in the store       demographics as linear features (e.g. + demo1 + demo2+.).** 
    ```{r}
      model.gm <- glm(formula=logmove ~ log(price)*brand*feat + AGE60 + EDUC + ETHNIC + INCOME +
      HHLARGE + WORKWOM + HVAL150 + SSTRDIST + SSTRVOL + CPDIST5 + CPWVOL5, data=oj)
    ```

    b. **What demographics are significantly (t>2) related to demand?**
    
    From what I see, all the demographics are pretty significant since the t values are all >2 or <-2
    ```{r}
      summary(model.gm);
    ```
    c. **How much did the adjusted R-squared improve with the addition of these variables?**
        
        the adjusted R-squared improved 9%;
    ```{r}
  model.lm <- lm(formula=logmove ~ log(price)*brand*feat  + AGE60 + EDUC + ETHNIC + INCOME + HHLARGE
                 + WORKWOM + HVAL150 + SSTRDIST + SSTRVOL + CPDIST5 + CPWVOL5, data=oj)
  model.old.lm <- lm(formula=logmove ~ log(price)*brand*feat, data=oj)
  
  model.lm.ars <- summary(model.lm)$adj.r.squared;
  model.old.lm.ars <- summary(model.old.lm)$adj.r.squared;
  
  (model.lm.ars-model.old.lm.ars)/model.old.lm.ars
    ```

2. Let's focus on two variables HVAL150 ("percent of HHs with homes >$150K") and one of your choosing. 

    a. **What are the means and percentiles of each of these variables? HINT: summary(oj$HVAL150) **
    ```{r}
    summary(oj$HVAL150);
    summary(oj$AGE60);
    ```

    b. **Using your coefficient estimates from the regression in 1b:**
        i. If we move from the median value of HVAL150 to the 75th percentile (3rd quartile), how much does log(quantity) change each week on average? HINT: using `coef(reg_output)["var_name"]` exports the coefficient on "var_name" from the regression model "reg_output". Similarly, `summary(df$var_name)` will output a bunch of summary statistics for the variable var_name in data frame df.  Using `summary(df$var_name)["3rd Qu."]` will take the level of the 3rd quantile from the summary of var_name.  Because we estimate things in logs you'll want to take the exponent of everything.
        ```{r}
        exp(coef(model.lm)["HVAL150"]*(summary(oj$HVAL150)["3rd Qu."]-summary(oj$HVAL150)["Median"]))
        ```
        
        ii.	If we move from the median value of AGE60 to the 75th percentile (3rd quartile), how much does log(quantity) change each week on average?
        ```{r}
          age60.sum <- summary(oj$AGE60);
          age60.q3 <- age60.sum["3rd Qu."]
          age60.median <- age60.sum["Median"];
          age60.e <- coef(model.lm)["AGE60"];
          exp(age60.e*(age60.q3-age60.median))
        ```

        iii. Base on this analysis, which is the more important predictor of demand?  
              age is a more important.
    c. **Now let's see if these variables impact price sensitivity. Add two interaction terms (with logprice) to the model to test this..  (Do this quickly.) **
        i. What are the coefficients on the interaction terms? 
        ```{r}
        model.b.gm <- glm(formula=logmove ~ log(price)*AGE60*HVAL150 +brand + feat  + EDUC + 
        ETHNIC + INCOME + HHLARGE+ WORKWOM + SSTRDIST + SSTRVOL + CPDIST5 + CPWVOL5,data=oj)
        model.b1.gm <- glm(formula=logmove ~ log(price)*AGE60,data=oj)
        model.b2.gm <- glm(formula=logmove ~ log(price)*HVAL150*AGE60,data=oj)
        
        
        ```
        ii.	Recall, positive values indicate lower price sensitivity and negative values indicate greater price sensitivity. Do your estimates make sense based on your intuition?
yes they do. since people who are older and/or have more expensive house are typically richer.
Therefore, they are less sensitive the price change/increase.

iii.	What are the coefficient estimates on the constants HVAL150 and your variable of choice? How do they compare to your regression from 1b?
```{r}

```

iv.	Similar to 2b, if we move from the median value of each variable to the 3rd quartile, how much does elasticity change? Based on this, which is more important to price sensitivity?

```{r}
set.seed(1);
Num_folds <-5;
model.formula <- logmove ~ log(price);

randomized <- sample(oj);
rowNumbers <- seq(1:nrow(oj));
partitioned <- rowNumbers %% Num_folds;
combined <- cbind(randomized, partitioned)

yterm <- toString(model.formula[2]); 
mse <- c(1:Num_folds);
for(i in 0:(Num_folds-1))
{
  test_sample<- randomized %>% filter(partitioned == i);
  training_sample <- randomized %>% filter(partitioned != i);
  model <- lm(training_sample,formula=model.formula);
  y_hat <- predict(model,newdata = test_sample);
  diff <- mean((y_hat - test_sample[[yterm]])^2);
  mse[i+1] <- diff; 
}
mean(mse)
```
