---
title: "NaiveBayes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("e1071")
library(e1071)
```

Creating example data

```{r}
train <- data.frame(class=c("spam","ham","ham","ham"), viagra=c("yes","no","no","yes"));

train
```
P(spam|viagra = yes) = P(viagra|spam) P(spam)/P(viagra)

p(viagra|spam) = 1

p(spam) = 1/4
p(viagra) = 2/4 = 1/2

(1*1/4) / (1/2)

2/4

1/2

syntax:
predict ~ given
```{r}
classifier <- naiveBayes(class ~ viagra, train)
classifier
```


estimated model

```{r}
test <- data.frame(viagra=c("yes"))
test$viagra <- factor(test$viagra, levels=c("no","yes"))
test

prediction <- predict(classifier, test, type="raw")
prediction
```

two predictors

```{r}
train$meet <- factor(c("yes","yes","yes","no"));

classifier <- naiveBayes(class ~ viagra + meet, train)
#p(viagra|spam) p(meet|spam)
classifier

#p(spam|viagra,meet) = p(viagra=yes,meet=yes|spam) P(spam)/p(viagra,meet)
#                    =            1                 1/4   /       1/4
#                    = 1
```

conditional property with two predictors

```{r}
test <- data.frame(viagra=c("yes"),meet=c("yes"))
test$viagra <- factor(test$viagra, levels=c("no","yes"))
test$meet <- factor(test$meet,levels=c("no","yes"))
test

prediction <- predict(classifier, test, type="raw")
prediction
```

Exercises

1. Suppose you have a database on four customers. You know their income and whether or not they bought your product. Create a data frame with this data.
```{r}
train <- data.frame(buy=c("yes","no","no","yes"),
                    income=c("high","high","medium","low"))
```
2. Using Bayes rule calculate the probability that a customer will buy your product given that he or she has high income.
p(buy|income=high) = p(income=high|buy) p(buy) / p(income=high)
= .5 * .5 / .5
= .5
3. Estimate naive Bayes model using your data above. What is the prior probability of someone buying your product? What is the probability that a customer has a high income give that he or she bought your product?

prior probability of someone buying the product is .5.
probability that a customer has a high income given that he/she has bought you product is .5;
```{r}
classifier <- naiveBayes(buy~income,train)
classifier
```
4. Using the model you estimated above predict the probability of buying given that a customer has high income. Is your result the same as the one you calculated by hand in question 1?

yap its the same
```{r}
test <- data.frame(income=c("high"))
test$income <- factor(test$income, levels=c("high","low","medium"))

prediction <- predict(classifier,test, type="raw")
prediction
```
5. Suppose you have a database on four customers. You know their gender, income and whether or not they bought your product. Create a data frame with this data.
```{r}
train$gender <- factor(c("male","female","female","male"))
```
6. Using Bayes rule calculate the probability that a customer will buy your product given that he has high income and male.

p(buy=yes|income=high,gender=male) = 1 (based on the table)
p(buy=yes|income=high,gender=male) = p(income=high,gender=male|buy=yes) * p(buy=yes) / p(income=high,gender=male)
=.5 * .5 / 1/4
= 1/4 / 1/4
= 1

7. Estimate naive Bayes model using your data above. What is the prior probability of someone buying your product? What is the probability that a customer has a high income given that he or she bought your product? What is the probability that a customer is male given that he bought your product?
```{r}
classifer <- naiveBayes(buy ~ income + gender, train)
classifer
```
8. Using the model you estimated above, predict the probability of buying given that a customer has a high income and is male. Is your result the same as the one you calculated by hand in question 1?

```{r}
test <- data.frame(income=c("high"),gender=c("male"))
test$income <- factor(test$income, levels=c("high","low","medium"))
test$gender <- factor(test$gender, levels=c("female","male"))
prediction <- predict(classifier, test, type="raw")
prediction
```
it is not the same as the my result.
